Analytic choices for analyzing multidimensional behavior - Many analysts test hypotheses about human speech.

### 2021-03-25 RR_manuscript/Draft_RR.tex
:memo: - edits to RR

Several other recent studies corroborated this analytic flexibility across different disciplines.
Both Dutilh et al. (2019) ad Starns et al. (2019) investigated analysts' choices when inferring theoretical constructs based on the same data set using computational models.
Both studies revealed vastly different modeling strategies, even though scientific conclusions were similar across analysis teams (see also Parker et al.~(accepted) on analytic flexibility in ecology and Botvinik-Nezer et al. 2020a on neuroimaging data).
Bastiaansen et al. (2020) crowdsourced clinical recommendations based on analyses of an individual patient.
Results suggests that analysts differed substantially regarding decisions related to both the statistical analysis of the data and the theoretical rationale behind interpreting the statistical results.

@@

When looking at phrase-level temporal domains, the number of possible analytic pipelines increases exponentially.
This plurality of analytic paths is illustrated in Figure \ref{fig:forkingPaths}.
Even if we have decided to compare fundamental frequency (f0) of only the word \emph{another} across the two utterances, there are still many choices to be made, all of which need to be justified.
For example, we could measure f0 at specific points in time like the onset of the window, the offset, or the midpoint.
We could also measure the value or time of the minimum or maximum f0 value.
We could summarize f0 across the entire window and extract the mean, median or standard deviation of f0, all of which have been used to analyze speech data (see Gordon and Roettger 2017).
And the garden of forking paths does not stop here.
Figure \ref{fig:forkingPaths}, illustrates a specific option to automatically calculate f0.
Knowing that the signals and the measures thereof can be noisy, another option would be to smooth them using one algorithm and degree over others, or remove values that substantially deviate from surrounding values or expectations, either manually or automatically, and so on.

@@

These decisions are usually made prior to any statistical analysis, but are at times revised a posteriori (i.e.~after data collection and/or preliminary analyses) in light of unforeseen or surprising outcomes.
These myriads of possible decisions are exponentiated by those researcher degrees of freedom related to statistical analysis (e.g. Wicherts et al. 2016).
In sum, speech data are complex physical signals that constitutes an as of yet unappreciated amount of analytic flexibility when choosing measures and operationalizing underlying theoretical constructs.
The present paper probes this garden of forking paths in the analysis of speech.
To assess the variability in data analysis pipelines across independent researchers, we will provide analytic teams with an experimentally elicited speech production data set.
The study aimed at investigating whether speakers acoustically modify utterances to signal unexpected referring expressions.
In the following, we will shortly introduce the research question of the original study.

@@

\begin{figure}
\includegraphics[width=1\linewidth]{/Users/ste/repos/many_analyses/figs/procedure_plot} \caption{Experimental procedure timeline.}\label{fig:procedural}
\end{figure}

@@

The two sentence prompts were used to create a focus contrast between the competitor and the target object.
If the competitor and target objects differed but their color did not (e.g.~\emph{yellow banana} vs \emph{yellow tomato}), the noun was discourse-pragmatically focused (the Noun Focus condition, NF).
If the objects were the same but differed in color (e.g.~\emph{yellow banana} vs \emph{blue banana}), the color adjective was discourse-pragmatically focused (the Adjective Focus condition, AF).
If both the color and the object differed (e.g.~\emph{yellow banana} vs \emph{blue tomato}), the whole noun phrase was discourse-pragmatically focused (the Adjective/Noun Focus condition, ANF).
The NF condition constituted the experimentally relevant condition, while the AF and ANF conditions acted as fillers.
The color-object combinations in the Noun Focus (NF) condition were manipulated with respect to their typicality.
The combinations were either typical (e.g.~\emph{orange mandarin}), medium typical (e.g.~\emph{greem tomato} ), or atypical (e.g.~\emph{yellow cherry}), as established by a norming study that was conducted prior to the production experiment (see Appendix XXX for details).

@@

We estimate that this process, from the time of an in-principle acceptance of this Stage 1 Registered Report to the end of Phase 5, will take XX months (Table X).
The factor most likely to delay our time line is the rate of completion of the original set of analyses by independent analysis teams.

@@

The initiating authors (SC, JC, TR) created an online landing page providing a general description of the project \url{https://many-speech-analyses.github.io} and a short pre-recorded slide show that summarizes the study and research question (LINK).
The project will be advertised via social media, using mailing lists for linguistic and psychological societies, and via word of mouth.
The target population comprises active speech science researchers with a graduate/doctoral degree (or currently studying for a graduate/doctoral degree) in relevant disciplines.
Researchers can choose to work independently or in a small team.
For the sake of simplicity, we will refer to single researcher or small teams as ``analysis teams.''
Recruitment for this project will commence upon receiving in-principle acceptance; we aim for a minimum of 12 analysis teams independently evaluating each data set (see sample size justification below).

We will simultaneously recruit volunteers to peer-review the analyses conducted by the other volunteers through the same channels.
Our goal is to recruit a similar number of peer-reviewers and analysts, and to ask each peer reviewer to review a minimum of four analyses.
If we are unable to recruit at least half the number of reviewers as analysis teams, we will ask analysts to serve also as reviewers (after they have completed their analyses).
All analysts and reviewers will share co-authorship on this manuscript and will participate in the collaborative process of producing the final manuscript.
All analysts will sign a consent (ethics) document (LINK).

Following Parker et al.~{[}ECO RR{]}, we identified a minimal sample size that would provide an estimate of heterogeneity of the effect sizes reported by the analysis teams (\(\tau^2\), i.e.~the variance \(\sigma^2_{\alpha_{\text{t}}}\), see \Cref{s:meta-est}).
In short, the minimal sample size is considered the number of analysis teams needed to allow an estimation of heterogeneity.
We consider a posterior distribution of \(\tau^2\), for which the 95\% credible interval (CrI) does not include 0 as sufficient.
The probability \(Pr\) that a CrI interval does not include 0 is obtained via the \emph{t}-statistic:

@@

The analysis teams will register and answer a demographic and expertise questionnaire (\texttt{intake\_form.pdf}).\footnote{PDF versions of all the questionnaires mentioned in this paper are available at \url{https://github.com/many-speech-analyses/many_analyses/tree/master/docs/questionnaires}.}
The questionnaire collects information on the analysts current position and self-estimated breadth and level of statistical expertise and acoustic analysis skills.
We will then request that they answer the question: \emph{Do speakers acoustically modify utterances to signal atypical word combinations?}
To do so, they will be given the data generated by the experiment described in \Cref{s:dataset}.
The data will include the audio recordings with corresponding time-aligned transcriptions in the form of Praat TextGrid files.\footnote{See \url{https://osf.io/rdtx5/?view_only=66989a3bbe034e188846e23f84913ffe}. Due to storage caps on the OSF, we will make the audio files available on \url{https://eudat.eu/catalogue/B2SHARE} in due course.}

@@

The posterior distribution of the population-level intercept \(\alpha\) will allow us to estimate the range of probable values of the standardized effect size \(\hat{\eta}\).
The posterior distribution will further allow us to investigate the effect of a set of analytic and demographic predictors, detailed in Section \ref{ana-factors}.
Crucially, the posterior distribution of \(\sigma_{\alpha_{\text{t}}}\) (the standard deviation of the group-level effect of team) will allow us to quantify the degree of variation between teams on a standardized scale.

@@

\hypertarget{ana-factors}{%
\subsubsection{Analytic and demographic predictors affecting effect sizes}\label{ana-factors}}

@@

\begin{itemize}
\tightlist
\item
  \emph{Measure of uniqueness} {[}numeric{]} of individual analyses for the set of predictors in each model (\(uniq_i\)).
\item
  \emph{Measure of conservativeness} {[}numeric{]} of the model specification, as the number of random/group-level effects included (\(cons_i\)).
\item
  \emph{Number of post-hoc changes to the acoustic measurements} {[}numeric{]} the teams will report to have carried out (\(phoc_i\)).
\item
  \emph{Number of models} {[}numeric{]} the teams will report to have run (\(modn_i\)).
\item
  \emph{Major dimension} {[}categorical{]} that has been measured to answer the research question (\(mdim_i\)).
\item
  \emph{Temporal window} {[}categorical{]} that the measurement is taken over (\(twin_i\)).
\item
  \emph{Average peer-review rating} {[}numeric{]}, as the mean of the peer-review ratings for each analysis.
\end{itemize}

@@

\textbf{Researcher-related factors}. We will include the following predictors about the teams:

@@

\textbf{Model specification}. The model will be fitted as measurement-error models, with the predictors detailed in the preceding paragraphs.
The outcome variable of the model will be the standardized effect sizes.

A normal distribution will be used as the likelihood function of \(\alpha_{t[i]}\).
The mean of \(\alpha_{t[i]}\) is modeled on the basis of the overall intercept \(\beta\) and on the coefficients \(\upsilon_n\) of each predictor.
The numeric predictors will be centered and scaled and the categorical predictors will be sum coded.
As the prior for \(i\) and \(\upsilon\) we will use a normal distribution with mean 0 and standard deviation 1.

@@

>>>>

Several other recent studies corroborated this analytic flexibility across different disciplines.
Both Dutilh et al. (2019) ad Starns et al. (2019) investigated analysts' choices when inferring theoretical constructs based on the same data set using computational models.

Both studies revealed vastly different modeling strategies, even though scientific conclusions were similar across analysis teams Botvinik-Nezer et al. (2020a).
Bastiaansen et al. (2020) crowdsourced clinical recommendations based on analyses of an individual patient.
Results suggests that analysts differed substantially regarding decisions related to both the statistical analysis of the data and the theoretical rationale behind interpreting the statistical results.

@@

When looking at phrase-level temporal domains, the number of possible analytic pipelines increases exponentially.
This plurality of analytic paths is illustrated in Figure \ref{fig:forkingPaths}.
Even if we have decided to compare fundamental frequency (f0) of only the word \emph{another} across the two utterances, there are still many choices to be made, all of which need to be justified.
For example, we could measure f0 at specific points in time like the onset of the window, the offset, or the midpoint.
We could also measure the value or time of the minimum or maximum f0 value.
We could summarize f0 across the entire window and extract the mean, median or standard deviation of f0, all of which have been used to analyze speech data (see Gordon and Roettger 2017).
And the garden of forking paths does not stop here.
Figure \ref{fig:forkingPaths}, illustrates one specific procedure to automatically calculate f0.
However, another options would be to filter the audio signal, to smooth the extracted f0 signal, to remove values that substantially deviate from surrounding values or expectations, either manually or automatically, and so on.

@@

These decisions are intended to be made prior to any statistical analysis, but are at times revised a posteriori (i.e.~after data collection and/or preliminary analyses) in light of unforeseen or surprising outcomes.
These myriads of possible decisions are exponentiated by those researcher degrees of freedom related to statistical analysis (e.g. Wicherts et al. 2016).
In sum, speech data are complex physical signals that constitutes an as of yet unappreciated amount of analytic flexibility when choosing measures and operationalizing underlying theoretical constructs.
The present paper probes this garden of forking paths in the analysis of speech.
To assess the variability in data analysis pipelines across independent researchers, we will provide analytic teams with an experimentally elicited speech production data set.
The study aimed at investigating whether speakers acoustically modify utterances to signal unexpected referring expressions.
In the following, we will shortly introduce the research question of the original study.

@@

\begin{figure}
\includegraphics[width=1\linewidth]{/Users/ste/repos/many_analyses/figs/procedure_plot} \caption{Experimental procedure - The upper row illustrates the trial sequence for the speaker (participant) and the lower row illustrates the trial sequence for the confederate. After a preview of 1500ms the speaker sees an arrow indicating one of the referent (b). Reading the orthographic instructions out loud, the speaker gives the confederate verbal instructions onto which referent they should drag the cube (c). The confederate, in turn, drags the black cube onto the target referent (d). Both the arrow and the orthographic instruction disappear from the speaker's screen and a new referent is indicated by an arrow on the same display alongside a new orthographic instruction (e). The speaker gives the confederate verbal instructions (f) which the confederate follows by dragging the cube onto the next referent (g).}\label{fig:procedural}
\end{figure}

@@

The two sentence prompts were used to create a focus contrast between the competitor and the target object.
If the competitor and target objects differed but their color did not (e.g.~\emph{yellow banana} vs \emph{yellow tomato}), the noun was discourse-pragmatically focused (the Noun Focus condition, NF).
If the objects were the same but differed in color (e.g.~\emph{yellow banana} vs \emph{blue banana}), the color adjective was discourse-pragmatically focused (the Adjective Focus condition, AF).
If both the color and the object differed (e.g.~\emph{yellow banana} vs \emph{blue tomato}), the whole noun phrase was discourse-pragmatically focused (the Adjective/Noun Focus condition, ANF).
The NF condition constituted the experimentally relevant condition, while the AF and ANF conditions acted as fillers.
The color-object combinations in the Noun Focus (NF) condition were manipulated with respect to their typicality.
The combinations were either typical (e.g.~\emph{orange mandarin}), medium typical (e.g.~\emph{greem tomato} ), or atypical (e.g.~\emph{yellow cherry}), as established by a norming study that was conducted prior to the production experiment (see HERE\footnote{Temporary link: \url{https://osf.io/5agn9/?view_only=3429acfd8eae4f0aafe2e17ad5b0983f}}).

@@

We estimate that this process, from the time of an in-principle acceptance of this Stage 1 Registered Report to the end of Phase 5, will take nine months.
The factor most likely to delay our time line is the rate of completion of the original set of analyses by independent analysis teams.

@@

The initiating authors (SC, JC, TR) created an online landing page providing a general description of the project and a short pre-recorded slide show that summarizes the study and research question (\url{https://many-speech-analyses.github.io}).
The project will be advertised via social media, using mailing lists for linguistic and psychological societies, and via word of mouth.
The target population comprises active speech science researchers with a graduate/doctoral degree (or currently studying for a graduate/doctoral degree) in relevant disciplines.
Researchers can choose to work independently or in a small team.
For the sake of simplicity, we will refer to single researcher or small teams as ``analysis teams.''
Recruitment for this project will commence upon receiving in-principle acceptance; we aim for a minimum of 12 analysis teams independently evaluating each data set (see sample size justification below).

We will simultaneously recruit volunteers to peer-review the analyses conducted by the other volunteers through the same channels.
Our goal is to recruit a similar number of peer-reviewers and analysts, and to ask each peer reviewer to review a minimum of four analyses.
If we are unable to recruit at least half the number of reviewers as analysis teams, we will ask analysts to serve also as reviewers (after they have completed their analyses).
All analysts and reviewers will share co-authorship on this manuscript and will participate in the collaborative process of producing the final manuscript.

Following Parker et al. (Parker et al. 2020), we identified a minimal sample size that would provide an estimate of heterogeneity of the effect sizes reported by the analysis teams (\(\tau^2\), i.e.~the variance \(\sigma^2_{\alpha_{\text{t}}}\), see \Cref{s:meta-est}).
In short, the minimal sample size is considered the number of analysis teams needed to allow an estimation of heterogeneity.
We consider a posterior distribution of \(\tau^2\), for which the 95\% credible interval (CrI) does not include 0 as sufficient.
The probability \(Pr\) that a CrI interval does not include 0 is obtained via the \emph{t}-statistic:

@@

The analysis teams will register and answer a demographic and expertise questionnaire (\texttt{intake\_form.pdf}).
PDF versions of all the questionnaires used in this study are available at LINK\footnote{Temporary link: \url{https://osf.io/h6z8w/?view_only=3457f2b01a5f4435b3d917eb7b5134e8}.}.
The questionnaire collects information on the analysts current position and self-estimated breadth and level of statistical expertise and acoustic analysis skills.
We will then request that they answer the question: \emph{Do speakers acoustically modify utterances to signal atypical word combinations?}
To do so, they will be given the data generated by the experiment described in \Cref{s:dataset}.
The data will include the audio recordings with corresponding time-aligned transcriptions in the form of Praat TextGrid files (LINK\footnote{Temporary link: \url{https://osf.io/5agn9/?view_only=3429acfd8eae4f0aafe2e17ad5b0983f}}).

@@

The posterior distribution of the population-level intercept \(\alpha\) will allow us to estimate the range of probable values of the standardized effect size \(\hat{\eta}\).
The posterior distribution will further allow us to investigate the effect of a set of analytic and demographic predictors, detailed in Section \ref{anares-preds}.
Crucially, the posterior distribution of \(\sigma_{\alpha_{\text{t}}}\) (the standard deviation of the group-level effect of team) will allow us to quantify the degree of variation between teams on a standardized scale.

@@

\hypertarget{anares-preds}{%
\subsubsection{Analytic and researcher-related predictors affecting effect sizes}\label{anares-preds}}

@@

\begin{itemize}
\tightlist
\item
  \emph{Measure of uniqueness} {[}numeric{]} of individual analyses for the set of predictors in each model (\(uniq_i\)).
\item
  \emph{Measure of conservativeness} {[}numeric{]} of the model specification, as the number of random/group-level effects included (\(cons_i\)).
\item
  \emph{Number of post-hoc changes to the acoustic measurements} {[}numeric{]} the teams will report to have carried out (\(phoc_i\)).
\item
  \emph{Number of models} {[}numeric{]} the teams will report to have run (\(modn_i\)).
\item
  \emph{Major dimension} {[}categorical{]} that has been measured to answer the research question (\(mdim_i\)).
\item
  \emph{Temporal window} {[}categorical{]} that the measurement is taken over (\(twin_i\)).
\item
  \emph{Average peer-review rating} {[}numeric{]}, as the mean of the overall peer-review ratings for each analysis.
\end{itemize}

@@

\textbf{Researcher-related factors}. We will also include the following predictors:

@@

Since we will obtain scores for each member in the analytic teams, we will calculate the team mean and standard deviation for each predictor.
These will be entered in the model formula as a measurement-error term (\texttt{me(mean,\ sd)} in brms).

\textbf{Model specification}. The model will be fitted as a measurement-error model, with the predictors detailed in the preceding paragraphs.
The outcome variable of the model will be the standardized effect sizes and related standard deviation.

A normal distribution will be used as the likelihood function of \(\alpha_{t[i]}\).
The mean of \(\alpha_{t[i]}\) is modeled on the basis of the overall intercept \(\beta\) and on the coefficients of each predictor.
The numeric predictors will be centered and scaled and the categorical predictors will be sum coded.
As the prior for the intercept and the predictors we will use a normal distribution with mean 0 and standard deviation 1.
The model will be run with the same settings as with the meta-analytic model.
The code used to run the model can be found here: INSERT LINK.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Parker2020}{}}%
Parker, Timothy, Hannah Fraser, Shinichi Nakagawa, Elise Bethany Gould, Simon Griffith, Peter Vesk, and Fiona Fidler. 2020. {``{Same data, different analysts: variation in effect sizes due to analytical decisions in ecology and evolutionary biology {[}passed peer review and granted in-principle acceptance March 2020{]}}.''} \emph{BMC Biology}. \url{https://doi.org/10.6084/m9.figshare.12034833.v1}.
